Isotropic Stellar Obliquities of Misaligned Planetary Systems

### 2022-12-24 src/tex/ms.tex
[showyourwork] overleaf sync

Many exoplanetary systems are found with high stellar obliquities. We build a hierarchical Bayesian framework to understand how stellar obliquities of misaligned systems are distributed. If stellar obliquities $\psi$ are isotropically distributed, the $\cos{\psi}$ distribution will be uniform, whereas if stellar obliquities have preferred angles, the $\cos{\psi}$ distribution will present concentrations.
We first apply our framework to \numistar systems with both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints. The $\cos{\psi}$ distribution is clustered near $-0.2$ ($\psi \sim 100\degr$), in agreement with the preponderance of perpendicular planets.
Interestingly, a similar $\cos{\psi}$ distribution could be inferred without using any $i_\star$ measurements, suggesting that the $\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
We then apply the framework to all \numall systems with sky-projected obliquity measurements. Unlike the observed-$i_\star$ sample, the $\cos{\psi}$ is  nearly uniformly distributed for misaligned systems, corresponding to an isotropic distribution of stellar obliquity.
We find the $\lambda$ distribution of the observed-$i_\star$ sample is clustered at $110\degr$, whereas the $\lambda$ distribution of the full sample is nearly uniform. The observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are often measured by the gravity-darkening method, which is biased toward detecting perpendicular planets.
We conclude that misaligned planetary systems, mostly Hot Jupiters, have isotropic stellar obliquities. When compared to theoretical predictions, the distribution is best consistency with the scattering of multiple giant planets after their convergent disk migration.
\end{abstract}

@@

The stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
For any given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$. Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$. If the stellar obliquity is not isotropically distributed but have some preferred angles, the $\cos{\psi}$ distribution will present concentrations at certain values.

In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky. Stellar obliquity $\psi$ can be computed as \citep[e.g.,][]{Fabrycky09}:
\begin{equation}\label{eqn:psi}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
Since the sky-projected stellar obliquities can be measured only in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.

To constrain the 3D stellar obliquity $\psi$, the measurements of the sky-projected obliquity $\lambda$ and stellar inclination $i_\star$ are required. The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology. For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ just from the sky-projected obliquity $\lambda$ assuming the stellar inclination $i_\star$ is isotropically distributed \citep[e.g.,][]{Fabrycky09}.

\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both sky-projected stellar obliquity and stellar inclination constraints. They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$. Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed, but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.

In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and show inferred stellar obliquity distributions. 
In Section~\ref{sec:interpretations}, we interpret the results. 
In Section~\ref{sec:discussion}, we discuss the implication of the stellar obliquity distribution for Hot Jupiter origins.

\section{3D Stellar Obliquity Distributions} \label{sec:hbm}
We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems. Figure~\ref{fig:graph} shows the probabilistic graphical model. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution. $\bm{\beta}$ is constrained by $N$ individual systems, where each system has $\psi_n$ that is calculated by the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$. For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$.
The parameter $\theta_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if the information is available.
Obs$_{\star,n}$ describes observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra. If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\theta_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}. If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves. If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$.

\begin{figure}
    \script{graph.py}
    \includegraphics[width=\linewidth]{figures/graph.pdf}
    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems. The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$. Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$. $\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period. Obs$_{\star,n}$ contains observations of the star, such as the stellar spectral and photometric and/or spectroscopic time series, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity.}
    \label{fig:graph}
\end{figure}

@@

\begin{figure*}
    \script{psi_dist.py}
    \begin{centering}
        \includegraphics{figures/psi_dist.pdf}
        \caption{The distributions of the cosine of 3D stellar obliquity $\cos{\psi}$. The left panel presents the inferred $\cos{\psi}$ distribution for the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements using the statistical model described in Figure~\ref{fig:graph}. The middle panel uses the same sample as the left panel but now removes the likelihood function on $i_\star$. The right panel presents the $\cos{\psi}$ distribution for all \numall systems that have sky-projected stellar obliquity $\lambda$ measurements and has no likelihood function on $i_\star$. The $\cos{\psi}$ distribution for the observed-$i_\star$ are similar, and both show a cluster near $\cos{\psi} = -0.2$, whereas the distribution for all systems does not have such a trend.}
        \label{fig:psi_dist}
    \end{centering}
\end{figure*}

@@

\begin{figure*}
    \script{coord_psi.py}
    \vspace*{-1cm}
    \gridline{
        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
    }
    \vspace*{-1.5cm}
    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.}
    \label{fig:coord}
\end{figure*}

@@

We find the underlying $\lambda$ distributions of the observed-$i_\star$ sample and the full sample by building another hierarchical Bayesian model. We again use a two-component mixture Beta distribution to describe well-aligned and misaligned systems. We extend the Beta distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$ and use the observed $\lambda$ and uncertainties for the likelihoods. Figure~\ref{fig:lam_dist} shows the $\lambda$ distributions of two samples. It clearly shows that in the observed-$i_\star$ sample, misaligned systems are clustered at $\sim 110\degr$, whereas in the full sample, misaligned systems are nearly uniformly distributed.

\begin{figure*}
    \script{lam_dist.py}
    \begin{centering}
        \includegraphics{figures/lam_dist.pdf}
        \caption{The distributions of sky-projected stellar obliquity ($\lambda$) of the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements (left panel) and all \numall systems that have sky-projected stellar obliquity measurements. The observed-$i_\star$ has a cluster of systems near $110\degr$, whereas such clustering is not found in the all-system distribution.}
        \label{fig:lam_dist}
    \end{centering}
\end{figure*}

@@

Furthermore, the measured-$i_\star$ sample could be biased towards perpendicular planets. Misaligned systems are often found around hot stars. However, to measure the inclinations of hot stars, the most prolific inclination measurement method, the rotational modulation method, is no longer applicable to hot stars due to their lack of starspots. Instead, the gravity-darkening method is often used. The method, however, could be biased since it only detects planets transiting across stellar disks with large brightness variations (i.e., planet crossing the the stellar poles, or perpendicular planets). In the measured-$i_\star$ sample, $6/21$ systems have stellar inclinations constrained via gravity darkening. All six planets are misaligned and have nearly polar orbits: 
KELT-9 ($\lambda = 85 \pm 0.2\degr$),
KELT-17 ($\lambda = 115.9 \pm 4.1 \degr$), 
Kepler-13 ($\lambda = 59.2 \pm 0.1 \degr$), 
MASCARA-1 ($\lambda = 69.5 \pm 3.0 \degr$),
MASCARA-4 ($\lambda = 112.5 \pm 1.6 \degr$), 
and WASP-189 ($\lambda = 89.3 \pm 1.4 \degr$).
The bias introduced by the sample is significant enough to produce the clustering in the $\lambda$ distribution of the observed-$i_\star$ sample.

\subsection{Hyperpriors}
We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}. We use different hyperprior distributions, including uniform distributions, log uniform distributions and truncated Normal distributions. In all hyperprior distributions, the clustering of perpendicular planets in observed-$i_\star$ sample is found and such clustering is much weaken in the full sample.

\section{Discussion} \label{sec:discussion}
The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution. When we apply our framework to all systems with $\lambda$ measurements, we find misaligned systems have nearly isotropic stellar obliquity distribution. We compare the inferred distribution from the observation to the prediction of mutual inclination of different eccentricity excitation mechanisms of Hot Jupiters. It is in best agreement with the outcome of scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.

>>>>

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, has important implications for the formation environment and evolution of a planetary system. In exoplanet observations, only sky-projected stellar obliquity $\lambda$ can be measured. To find the stellar obliquity, information on the stellar inclination $i_\star$ is required.
Here we show that, while it is important to measure the stellar inclination to tightly constrain the stellar obliquity of an individual system, stellar inclination is not necessarily required to infer the population stellar obliquity distribution. The stellar obliquity distribution can be inferred purely from the projected stellar obliquities, or even from the stellar inclinations.
We introduce a flexible, hierarchical Bayesian framework to infer stellar obliquity distribution of exoplanetary systems that can be applied to a sample with measured projected stellar obliquities, stellar inclinations, or both.
We apply the framework to all systems with measured projected stellar obliquities, which are mostly hot Jupiters, and find that the stellar obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquity distribution. When compared to hot Jupiter stellar obliquity predictions, the distribution is most consistent with the outcome of the scattering of multiple giant planets after their convergent disk migration.

\end{abstract}

@@

Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. The angle has implications for the formation environment of a planetary system and its consequently dynamical evolution. Exoplanetary systems present a wide range of stellar obliquities.

In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky. Stellar obliquity $\psi$ can be computed from $\lambda$ as \citep[e.g.,][]{Fabrycky09}:
\begin{equation}\label{eqn:psi}
    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
\end{equation}
where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.

To constrain the 3D stellar obliquity $\psi$, the measurements of the sky-projected obliquity $\lambda$ and stellar inclination $i_\star$ are required. The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology. Since the stellar inclination measurement is degenerate between $i_\star$ and $\pi - i_\star$, we assume $i_\star$ is less than $90\degr$.

For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ just from the sky-projected obliquity $\lambda$, assuming the stellar inclination $i_\star$ is isotropically distributed.

In this work, we find an approach to infer $\psi$ distribution from $\lambda$ distribution.

\begin{figure*}[ht!]
    \script{coord_psi.py}
    \gridline{
        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
    }
    \vspace*{-1.5cm}
    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.}
    \label{fig:coord}
\end{figure*}

In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
In Section~\ref{sec:jacobian}
In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and show inferred stellar obliquity distributions. 
In Section~\ref{sec:application}, we apply our framework to observed exoplanetary systems.

\section{Transformation between the $\psi$ and $\lambda$ distributions}\label{sec:jacobian}

For any given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$. Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$. If the stellar obliquity instead has some preferred angles, the $\cos{\psi}$ distribution will present concentrations at certain values.

Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly/isotropically distributed. We first find relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$, where $\theta$ is the azimuthal angle of the stellar spin axis, as shown in Figure~\ref{fig:coord}. We transform $\bf{n}_{\star}$ in two coordinate systems to Cartesian, pair their terms, and get:
\begin{align}
    \sin{\psi}\cos{\theta} = \cos{i_\star}& \label{eq:coord1}\\
    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \label{eq:coord2}\\
    \cos{\psi} = \cos{\lambda}\sin{i_\star} \label{eq:coord3}&.
\end{align}
Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \tan[-1](\tan{\psi}\sin{\theta})$.

First, we find the $\lambda$ distribution for a given $\cos{\psi}$ distribution. If the stellar obliquity is isotropically distributed, the $\cos{\psi}$ distribution will be uniform, and the $\lambda$ distribution will also be uniform. If the stellar obliquity has preferred angles, i.e., the $\cos{\psi}$ distribution will have concentrations, what will be the corresponding $\lambda$ distribution? We first find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$. Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) with Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$. Reorganize the equation, and we also get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$.
The distribution transformation follows
\begin{equation}
    p_{\cos{\lambda}} = 2 \int_{0}^{1} \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p_{\cos{\psi}} p_{\cos{\theta}} d\cos{\theta}.
\end{equation}
We have $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$. Assuming $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$\footnote{The $\theta$ is defined between $-\pi/2$ and $\pi/2$ since we limit $0 \leq i_\star \leq \pi/2$.}, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$. Combining all the parts together, we get
\begin{equation}\label{eqn:jac}
    p_{\cos{\lambda}} = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p_{\cos{\psi}} d\cos{\theta}.
\end{equation}

If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral derives $p_{\cos{\lambda}} = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda = 1/\pi$. The $\lambda$ is uniformly distributed. If $\cos{\psi}$ follows other distributions, we may integrate Equation~(\ref{eqn:jac}) numerically, find the $\cos{\lambda}$ distribution, and solve for the $\lambda$ distribution.

Similarly, we could derive the distribution of $\cos{i_\star}$ for a given $\cos{\psi}$ distribution via the Jacobian transformation and marginalizing over $\cos{\theta}$. We begin with
\begin{equation}
    p_{\cos{i_\star}} = 4 \int_{0}^{1} \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p_{\sin{\psi}} p_{\cos{\theta}} d\cos{\theta},
\end{equation}
and from Equation(\ref{eq:coord1}), we find $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$. Similarly, we assume $\theta$ is uniformly distributed, and $p_{\cos{\theta}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$. Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = p_{\cos{\psi}}\sqrt{1-\cos^2{\psi}}/\cos{\psi}$. Combining all the pieces together, we get
\begin{equation}
    p_{\cos{i_\star}} = \frac{4}{\pi} \int_{\cos{i_\star}}^{1} \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \frac{1}{\sqrt{1-\cos^2{\theta}}} p_{\cos{\psi}} d\cos{\theta}.
\end{equation}
If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ is uniform.

We note that we could not infer $\psi$ from $\lambda$ and $\theta$ since they are not independent variables. Instead, we will have to infer $\psi$ from $\lambda$ and $i_\star$.

\begin{figure*}[ht!]
    \script{transform.py}
    \includegraphics[width=\linewidth]{figures/transform.pdf}
    \caption{Distribution transformation.}
    \label{fig:transform}
\end{figure*}

\section{Hierarchical Bayesian Framework} \label{sec:hbm}

\begin{figure}[ht!]
    \script{graph.py}
    \includegraphics[width=\linewidth]{figures/graph.pdf}
    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems. The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$. Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$. $\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period. Obs$_{\star,n}$ contains observations of the star, such as the stellar spectral and photometric and/or spectroscopic time series, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity.}
    \label{fig:graph}
\end{figure}

We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems. Figure~\ref{fig:graph} shows the probabilistic graphical model. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution. $\bm{\beta}$ is constrained by $N$ individual systems, where each system has $\psi_n$ that is calculated by the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$. For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$.
The parameter $\theta_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if the information is available.
Obs$_{\star,n}$ describes observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra. If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\theta_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}. If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves. If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$.

@@

\begin{figure*}[ht!]
    \script{psi_dist.py}
    \begin{centering}
        \includegraphics{figures/psi_dist.pdf}
        \caption{The distributions of the cosine of 3D stellar obliquity $\cos{\psi}$. The left panel presents the inferred $\cos{\psi}$ distribution for the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements using the statistical model described in Figure~\ref{fig:graph}. The middle panel uses the same sample as the left panel but now removes the likelihood function on $i_\star$. The right panel presents the $\cos{\psi}$ distribution for all \numall systems that have sky-projected stellar obliquity $\lambda$ measurements and has no likelihood function on $i_\star$. The $\cos{\psi}$ distribution for the observed-$i_\star$ are similar, and both show a cluster near $\cos{\psi} = -0.2$, whereas the distribution for all systems does not have such a trend.}
        \label{fig:psi_dist}
    \end{centering}
\end{figure*}

@@

@@

We find the underlying $\lambda$ distributions of the observed-$i_\star$ sample and the full sample by building another hierarchical Bayesian model. We again use a two-component mixture Beta distribution to describe well-aligned and misaligned systems. We extend the Beta distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$. We use the observed $\lambda$ and uncertainties for the likelihoods. Figure~\ref{fig:lam_dist} shows the $\lambda$ distributions of two samples. It clearly shows that in the observed-$i_\star$ sample, misaligned systems are clustered at $\sim 110\degr$, whereas in the full sample, misaligned systems are nearly uniformly distributed.

\begin{figure*}[ht!]
    \script{lam_dist.py}
    \begin{centering}
        \includegraphics{figures/lam_dist.pdf}
        \caption{The distributions of sky-projected stellar obliquity ($\lambda$) of the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements (left panel) and all \numall systems that have sky-projected stellar obliquity measurements. The observed-$i_\star$ has a cluster of systems near $110\degr$, whereas such clustering is not found in the all-system distribution.}
        \label{fig:lam_dist}
    \end{centering}
\end{figure*}

@@

Furthermore, the measured-$i_\star$ sample could be biased towards perpendicular planets. Misaligned systems are often found around hot stars. However, to measure the inclinations of hot stars, the most prolific inclination measurement method, the rotational modulation method, is no longer applicable to hot stars due to their lack of starspots. Instead, the gravity-darkening method is often used. The method, however, could be biased since it only detects planets transiting across stellar disks with large brightness variations (i.e., planet crossing the stellar poles or perpendicular planets). In the measured-$i_\star$ sample, $6/21$ systems have stellar inclinations constrained via gravity darkening. All six planets are misaligned and have nearly polar orbits: 
KELT-9 ($\lambda = 85 \pm 0.2\degr$),
KELT-17 ($\lambda = 115.9 \pm 4.1 \degr$), 
Kepler-13 ($\lambda = 59.2 \pm 0.1 \degr$), 
MASCARA-1 ($\lambda = 69.5 \pm 3.0 \degr$),
MASCARA-4 ($\lambda = 112.5 \pm 1.6 \degr$), 
and WASP-189 ($\lambda = 89.3 \pm 1.4 \degr$).
The bias introduced by the sample is significant enough to produce the clustering in the $\lambda$ distribution of the observed-$i_\star$ sample.

\subsection{Hyperpriors}
We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}. We use different hyperprior distributions, including uniform distributions, log uniform distributions, and truncated Normal distributions. In all hyperprior distributions, the clustering of perpendicular planets in the observed-$i_\star$ sample is found, and such clustering is much weakened in the full sample.

\section{Discussion} \label{sec:discussion}

\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both sky-projected stellar obliquity and stellar inclination constraints. They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$. Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.

The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution. When we apply our framework to all systems with $\lambda$ measurements, we find misaligned systems have nearly isotropic stellar obliquity distribution. We compare the inferred distribution from the observation to the prediction of mutual inclination of different eccentricity excitation mechanisms of Hot Jupiters. It is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.
